---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
raw_data <- read.csv("activity.csv")
intervals<-levels(factor(raw_data$interval))
dates <-levels(factor(raw_data$date))
mmm <- matrix(raw_data$steps, nrow = length(dates), ncol = length(intervals), dimnames = list(dates,intervals),byrow = TRUE)

```

## What is mean total number of steps taken per day?
 `total_step` is a variable to store total number of steps for each date
```{r}
total_step<-rowSums(mmm, na.rm = TRUE)
hist(total_step, xlab = "Total steps per day", main = "Total steps per day")
mean(total_step)
median(total_step)
```

## What is the average daily activity pattern?
`means_activity` is a variable to store average date of steps for each interval.
Output is time interval in format HMM
```{r}
means_activity <-colMeans(mmm,na.rm = TRUE)
plot(dimnames(mmm)[[2]],means_activity,type = 'l',xlab = "time interval", main = "average activity", ylab = 'number of step')
names(which.max(means_activity))
```
## Imputing missing values
Total number of missing values in the dataset
```{r}
summary(raw_data$steps)[7]
```
A strategy for filling in all of the missing values in the dataset:
I use mean for that 5-minute interval or 0 if NA. This values was found on previous step and store in `means_activity`
```{r}
new_mmm<-mmm
for ( i in 1:nrow(new_mmm)){
  new_mmm[i,][is.na(mmm[i,])]<- means_activity[is.na(mmm[i,])]
}
total_step_new <-rowSums(new_mmm, na.rm = TRUE)
hist(total_step_new, xlab = "Total steps per day", main = "Total steps per day after filling NA's")
mean(total_step_new)
median(total_step_new)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
sortWeekdays <- function(day){
  retv <- 0
  day <- as.Date(day)
  day<- weekdays(day)
  if (day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")) {
    retv<-"Weekday"
  }
  if (day %in% c("Sunday","Saturday")) {
    retv<-"Weekend"
  }
  retv
} 

#week_factor<- factor(sapply(as.vector(dimnames(new_mmm)[1]),sortWeekdays))

weekdy <- dates
for ( i in 1:length(weekdy)) {
  weekdy[i] <- sortWeekdays(weekdy[i])
  }
means_activity_wd <-colMeans(new_mmm[weekdy == "Weekday",],na.rm = TRUE)
plot(dimnames(new_mmm)[[2]],means_activity_wd,type = 'l',xlab = "time interval", main = "average activity weekday", ylab = 'number of step')
means_activity_we <-colMeans(new_mmm[weekdy == "Weekend",],na.rm = TRUE)
plot(dimnames(new_mmm)[[2]],means_activity_we,type = 'l',xlab = "time interval", main = "average activity weekend", ylab = 'number of step')

```
